### 2020-09-10 manuscript.Rmd
Initial Rmd

---
title: "INSERT_TITLE_HERE"
author: 
  - name: Philip Parker
    affiliation: Australian Catholic University

date: September 8, 2020
output: 
  hrbrthemes::ipsum:
    toc: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 300)
```

## References