### 2020-11-02 supplementary_materials.Rmd
Add supp material
Did the following:

* Added `supplementary_materials.Rmd`
* Modified function to produce full model outputs
* Added function to knit both manuscript and supp materials
* Updated pipeline with new functions
* Removed all old versions of codebook that have names in them

Should close #5

---
title: "Supplementary Materials: School Socioeconomic Status Context and Social Skills in Children"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
library(targets)
library(scales)
library(tidyverse)
library(DT)

```

# Full Linear Model

```{r}
tar_load(supp_linear)
supp_linear %>% 
  datatable(filter = 'top', caption = "Table S1. Full Linear Model Outputs")
```

# Full Interaction Model

```{r}
tar_load(supp_interaction)
supp_interaction %>%
  datatable(filter = 'top', caption = "Table S2. Full Interaction Model Outputs")
```

# Codebook

```{r, child="data/social_skills_codebook.Rmd"}

```