### 2020-11-06 supplementary_materials.Rmd
Ed Researcher Version

---
title: "Supplementary Materials: School Socioeconomic Status Context and Social Skills in Children"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)
library(targets)
library(scales)
library(tidyverse)
library(DT)
library(knitr)
```

# Participant Flow Diagram

```{r}
library(magick)
pdf = magick::image_read(here::here("fig","flow.pdf"))
image_convert(pdf, "png")
```

# Full Linear Model

```{r}
tar_load(supp_linear)
supp_linear %>% 
  datatable(filter = 'top', caption = "Table S1. Full Linear Model Outputs") %>%
    formatRound(columns=4:6, digits=3)
```

# Full Interaction Model

```{r}
tar_load(supp_interaction)
supp_interaction %>%
  datatable(filter = 'top', caption = "Table S2. Full Interaction Model Outputs") %>%
    formatRound(columns=4:6, digits=3)
```

# Codebook

```{r, child="data/social_skills_codebook.Rmd"}
```